abxD01 Paper

### 2015-03-07 PAPER.Rmd
Edit Methods section

The microbiome, or the diverse community of microorganisms living in and on the body, has an integral role in deterring pathogen colonization and infection. This native protection by the microbiome from invasive pathogenic species is termed colonization resistance. It is well established that the gut bacterial microbiome is critical in the hosts’ defense against the pathogen Clostridium difficile. When this indigenous community is perturbed, this often leads to a loss of resistance. This is especially important in many hospital settings where patients are not only exposed to various types and degrees of perturbations, such as antibiotics, diet changes, chemotherapy, but also exposed to the environmental *C. difficile* spores. Among hospital-acquired infections in the United States, *C. difficile* infections (CDI) are the number one reported and are responsible for 14,000 deaths a year. It is not completely understood how different perturbations affect subsequent colonization resistance, nor do we have a comprehensive view of the taxonomic changes associated with a loss in colonization resistance.

There is a need for more knowledge about interactions between members of the community. Know that fmts work [anna’s paper], but don’t know why, except that you recover the diversity and are receiving an established community already that works. Individual probiotics have not worked with complete success [dendukuri 2005], whereas studies find that combinations of bacteria often work better than by themselves [lawley, pamer, stein, hickson 2007]. It’s also common for bacteria to change their behavior in the context of other bacterial species [Crost 2010, quorum sensing-biofilms, recognizing fellow bacterial species vs killing others, synergistic effects (sieuwerts 2010)]. Understanding the scope of relationships between bacteria in a community is necessary to learn about its response, in concert with its host, to perturbations including antibiotic-induced, pathogen invasion, probiotics doses, and diet changes, etc.

@@

Though we are limited in a resolution at the strain and species level, we have identified groups of related species with antagonistic relationships with *C. difficile* colonization. We identified several models with equally robust quality statistics and observed that across the top models a subset of 5 species were consistently present. These findings __

@@

***Animal Care.*** We used 5-8 week old C57Bl/6 mice for all our experiments. These mice were housed in our SPF room in the animal facility at the University of Michigan. All animal-related protocols and experiments were approved by the University Committee on Use and Care of Animals at the University of Michigan.

***Antibiotic Administration.*** Mice were administered one of seven different antibiotics, including cefoperazone, vancomycin, metronidazole, streptomycin, ciprofloxacin, ampicillin, and clindamycin. The route of administration depended on the specific antibiotic. Cefoperazone (either 0.5mg/ml, 0.3mg/ml, or 0.1mg/ml), vancomycin (0.625mg/ml, 0.3mg/ml, or 0.1mg/ml), streptomycin (5mg/ml, 0.5mg/ml, or 0.1mg/ml), metronidazole (0.5mg/ml), and ampicillin (0.5mg/ml) were all administered in the mouse drinking water for 5 days. Ciprofloxacin (10 mg/kg) was administered via oral gavage, and clindamycin (10 mg/kg) was administered via IP injection. All antibiotic treated mice were given 24 hours of recovery on regular drinking water prior to *C. difficile* challenge.

***C. difficile Preparation and Challenge*** In all our experiments we used *C. difficile* strain 630Δerm [cite dena lyras]. The spores used were prepared from a single large batch made following previously established protocols [cite paul?]. Spores were quantified within the week prior to *C. difficile* challenge for each experiment [established protocol I can cite?]. Spores were quantified by boiling a small volume at 65°C for 20 minutes to kill off residual vegetative cells. The volume was then diluted and plated on TCCFA plates. These were incubated at 37°C anaerobically overnight. The next day *C. difficile* colony forming units (CFU) were counted and the concentration of spores in the working stock was back calculated. On the day of challenge, *C. difficile* spores, at a final concentration of 103, were administered to mice via oral gavage. Immediately following this challenge, the remaining *C. difficile* inoculum was diluted in a series and plated to confirm the correct dosage.

***Sample Collection and Plating*** Fecal samples were freshly collected for each mouse immediately prior to the administration of any antibiotic, during any antibiotic recovery days, and days 0 and 1 post *C. difficile* challenge. Each fecal sample on day 1 was weighed and diluted under anaerobic conditions with anaerobic PBS 1:10 based on pellet weight. A dilution series was then plated onto TCCFA plates [cite plates] and incubated at 37°C anaerobically overnight. The colony forming units (CFU) were counted and calculated the subsequent day.

***DNA Extraction and Sequencing*** Total bacterial DNA was extracted from each untreated and day 0 stool sample using the MOBIO PowerSoil®-htp 96 Well Soil DNA Isolation Kit. Sequencing preparation and sequencing was carried out as previously described (Kozich JJ et al. 2013). Briefly, the V4 region of the 16S rRNA gene was amplified, purified, and pooled in equimolar concentrations. These libraries were then mixed with Illumina-prepared PhiX control libraries. Sequencing was performed using Illumina’s MiSeq platform to produce paired 250-nt reads. Mock/error/error of run?

***Sequence Curation*** These sequences were curated using mothur as previously described (Kozich JJ et al and Schloss PD et al). The full pipeline is also demonstrated online at the mother website (http://www.mothur.org/wiki/MiSeq_SOP). Sequences were binned into operational taxonomic units (OTUs) using a 3% dissimilarity cutoff. Taxonomic assignments were determined by using a naïve Bayesian classifier with the Ribosomal Database Project (RDP) training set (version 9) requiring an 80% bootstrap confidence score. In order to avoid biases due to uneven sampling, samples were normalized to 1,625 sequences per samples [Schloss, reducing effects]. For those samples with more than 1,625 sequences per samples, the range was __, the mean was __, the median was __, the median absolute deviation was __. All 16S rRNA gene sequence data and the MIMARKS table are available at __.

***Statistical Analysis & Modeling*** Across all data, we set a relative abundance threshold of an average of 1% in at least one experimental group. The 44 OTUs that met this cutoff were included in the training of the model. We used the "randomForest" package in R to build a regression-based random forest model to predict *C. difficile* colonization levels on day 1 based on the microbiome data from day 0. All statistical analyses were conducted using R version 3.1.2.

>>>>

The microbiome, or the diverse community of microorganisms living in and on the body, has an integral role in deterring pathogen colonization and infection [ref]. This native protection by the microbiome from invasive pathogenic species is termed colonization resistance [ref]. It is well established that the gut bacterial microbiome is critical in the hosts’ defense against the pathogen Clostridium difficile [ref]. When this indigenous community is perturbed, this often leads to a loss of resistance. This is especially important in many hospital settings where patients are not only exposed to various types and degrees of perturbations, such as antibiotics, diet changes, chemotherapy, but also exposed to the environmental *C. difficile* spores [ref]. Among hospital-acquired infections in the United States, *C. difficile* infections (CDI) are the number one reported and are responsible for 14,000 deaths a year [ref]. It is not completely understood how different perturbations to the taxonomic composition of the gut microbiota result in a loss of colonization resistance.

<not done>
There is a need for more knowledge about interactions between members of the community. Know that fmts work [anna’s paper], but don’t know why, except that you recover the diversity and are receiving an established community already that works. Individual probiotics have not worked with complete success [dendukuri 2005], whereas studies find that combinations of bacteria often work better than by themselves [lawley, pamer, stein, hickson 2007]. It’s also common for bacteria to change their behavior in the context of other bacterial species [Crost 2010, quorum sensing-biofilms, recognizing fellow bacterial species vs killing others, synergistic effects (sieuwerts 2010)]. Understanding the scope of relationships between bacteria in a community is necessary to learn about its response, in concert with its host, to perturbations including antibiotic-induced, pathogen invasion, probiotics doses, and diet changes, etc.

@@

Though we are limited in a resolution at the strain and species level, we have identified groups of related species with antagonistic relationships with *C. difficile* colonization. We identified several models with equally robust quality statistics and observed that across the top models a subset of 5 species were consistently present. These findings __
</not done>

@@

***Animal Care.*** We used 5-8 week old C57Bl/6 mice for all our experiments. These mice were reared under SPF conditions within the animal facility at the University of Michigan. All animal-related protocols and experiments were approved by the University Committee on Use and Care of Animals at the University of Michigan.

***Antibiotic Administration.*** Mice were administered one of seven different antibiotics, including cefoperazone, vancomycin, metronidazole, streptomycin, ciprofloxacin, ampicillin, and clindamycin. The route of administration depended on the specific antibiotic. Cefoperazone (either 0.5 mg/ml, 0.3mg/ml, or 0.1 mg/ml), vancomycin (0.625 mg/ml, 0.3 mg/ml, or 0.1 mg/ml), streptomycin (5 mg/ml, 0.5 mg/ml, or 0.1 mg/ml), metronidazole (0.5 mg/ml), and ampicillin (0.5 mg/ml) were all administered in the mouse drinking water for 5 days. Ciprofloxacin (10 mg/kg) was administered via oral gavage and clindamycin (10 mg/kg) was administered via IP injection. Mice that did not receive antibiotics were used as negative controls for these experiments.

***C. difficile Preparation and Challenge*** All antibiotic-treated mice were given 24 hours to recover with untreated drinking water prior to *C. difficile* challenge. *C. difficile* strain 630Δerm spores were used in all experiments [cite dena lyras]. Spores were prepared from a single large batch whose concentration was determined within the week prior to each *C. difficile* challenge [cite paul?]. On the day of challenge 10^3 *C. difficile* spores were administered to mice via oral gavage. Immediately following this challenge, the remaining *C. difficile* inoculum was diluted in a series and plated to confirm the correct dosage.

***Sample Collection and Plating*** Fecal samples were freshly collected for each mouse on the day of *C. difficile* challenge (day 0). On the day after the challenge (day 1), another fecal sample was weighed and diluted under anaerobic conditions with anaerobic PBS. The number of colony forming units (CFU) were counted following 24 hrs growth on TCCFA plates at 37°C under anaerobic conditions [cite plates].

***DNA Extraction and Sequencing*** Total bacterial DNA was extracted from each  day 0 stool sample using the MOBIO PowerSoil®-htp 96 Well Soil DNA Isolation Kit. Sequencing preparation and sequencing was carried out as previously described (Kozich JJ et al. 2013). Briefly, the V4 region of the 16S rRNA gene was amplified, purified, and pooled in equimolar concentrations. These libraries were then mixed with Illumina-prepared PhiX control libraries.

***Sequence Curation*** These sequences were curated using mothur as previously described (Kozich JJ et al and Schloss PD et al). Sequences were binned into operational taxonomic units (OTUs) using a 3% dissimilarity cutoff. Taxonomic assignments were determined by using a naïve Bayesian classifier with the Ribosomal Database Project (RDP) training set (version 9) requiring an 80% bootstrap confidence score. In parallel to the fecal samples, we also sequeced a mock community where we knew the true sequence of the 16S rRNA gene sequences. Analysis of the mock community data indicated that the error rate following our cuation procedure was 0.0X%. In order to avoid biases due to uneven sampling, samples were normalized to 1,625 sequences per samples [Schloss, reducing effects]. All 16S rRNA gene sequence data and metadata are available through the
Sequence Read Archive under accession XXXXXX.

***Statistical Analysis & Modeling*** Random forest regression models were constructed using the randomForest R package using XX,000 trees [ref]. The
model was constructed using the relative abundance of those OTUs that had an
average relative abundance within any antibiotic treatment group of 1%. This
resulted in 44 OTUs. The regression was performed using the log10 transformation
of the number of CFU per gram of fecal material as the dependent variable. All analyses were conducted using R version 3.1.2. Complete analysis scripts are
available at the online repository for this study (https://github.com/SchlossLab/Schubert_AbxD01_mBio_2015).
